---
title: "Proportion of red beads"
author: "Faisal Jan"
format: html
execute: 
  echo: false
---

```{r}
#| message: false
library(tidyverse)
library(tidymodels)
library(broom)
library(knitr)
library(marginaleffects)
```

```{r}
#| cache: true

urn <- tibble(color = c(rep("red", 400), rep("white", 600))) |>
  mutate(color = as.factor(color))

shovel <- slice_sample(urn, n = 50)

fit_color <- logistic_reg(engine = "glm") |> fit(color ~ 1, data = shovel)
```

```{r}
fit_color |> tidy(conf.int = TRUE) |>
  mutate(
    estimate = round(estimate, 3),
    std.error = round(std.error, 3),
    statistic = round(statistic, 3),
    p.value = round(p.value, 3),
    conf.low = round(conf.low, 3),
    conf.high = round(conf.high, 3)
  ) |>
  select(
    Term = term,
    Estimate = estimate,
    `Std. Error` = std.error,
    `z-Statistic` = statistic,
    `p-value` = p.value,
    `CI Low` = conf.low,
    `CI High` = conf.high
  ) |>
  kable(
    caption = "Logistic Regression Coefficient Summary",
    digits = 3,
    align = "c"
  )
```

```{r}
plot_predictions(fit_color, type = "prob", condition = "group") +
  scale_fill_manual(values = c("red" = "#D73027", "white" = "#F0E442")) +
  labs(
    title = "Predicted Probability by Bead Color",
    subtitle = "Red group has slightly higher predicted probability than White",
    x = "Bead Color",
    y = "Predicted Probability",
    caption = "Source: Output from plot_predictions(fit_color, type = 'prob')"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold"),
    plot.subtitle = element_text(margin = margin(b = 10)),
    axis.title.y = element_text(margin = margin(r = 10)),
    axis.title.x = element_text(margin = margin(t = 10))
  ) +
  ylim(0, 0.7)
```

From a lot of beads in a urn, it will probably be costly to find the proportion of a specific color of beads. Using 1,000 beads of color white and red in a simulated urn, I seek to predict what proportion of beads are red. Collecting samples from the population may contains bias, so it is important to mix the urn before taking samples. I model the probability of bead being red, is a binary outcome variable that has two values; red and white, as a logistic function based on the observed distribution in the urn or in shovel samples. Red beads are predicted at around 46% in proportion of overall beads in urn, with a 95% confidence interval ranging from 21% to 47%.

$$
Y = \beta_0 + \beta_1 X_1 + \beta_2 X_2 + \cdots + \beta_n X_n + \varepsilon
$$

$$
\hat{Y} = \hat{\text{color}} = \frac{1}{1 + \exp\left(0.663 + \beta_1 X_1 + \beta_2 X_2 + \cdots + \beta_n X_n + \varepsilon\right)}
$$
